---
title: "Home Work - Working with Cardio Dataset"
author: "Ashutosh Shirsat"
date: \today
format: 
   pdf:
    toc: true
    toc-title: Contents
---

{{< pagebreak >}}

# Data set Description

Dataset "*data_cardio.csv*" has the following columns i.e. variables :

+--------------+--------------------------+--------------+----------------------+
| Nr.          | Variable Name            | Column Name  | Scale                |
+==============+==========================+==============+======================+
| 1            | Age                      | age          | int (days)           |
+--------------+--------------------------+--------------+----------------------+
| 2            | Height                   | height       | int (cm)             |
+--------------+--------------------------+--------------+----------------------+
| 3            | Weight                   | weight       | float (kg)           |
+--------------+--------------------------+--------------+----------------------+
| 4            | Gender                   | gender       | Categorical code     |
|              |                          |              |                      |
|              |                          |              | 1 : male             |
|              |                          |              |                      |
|              |                          |              | 0 : female           |
+--------------+--------------------------+--------------+----------------------+
| 5            | Systolic blood pressure  | ap_hi        | int                  |
+--------------+--------------------------+--------------+----------------------+
| 6            | Diastolic blood pressure | ap_lo        | int                  |
+--------------+--------------------------+--------------+----------------------+
| 7            | Cholesterol              | cholesterol  | Categorical code     |
|              |                          |              |                      |
|              |                          |              | 1: normal            |
|              |                          |              |                      |
|              |                          |              | 2: above normal      |
|              |                          |              |                      |
|              |                          |              | 3: well above normal |
+--------------+--------------------------+--------------+----------------------+
| 8            | Glucose                  | gluc         | Categorical code     |
|              |                          |              |                      |
|              |                          |              | 1: normal            |
|              |                          |              |                      |
|              |                          |              | 2: above normal      |
|              |                          |              |                      |
|              |                          |              | 3: well above normal |
+--------------+--------------------------+--------------+----------------------+
| 9            | Smoking                  | smoke        | Categorical -Binary  |
|              |                          |              |                      |
|              |                          |              | 0 : No               |
|              |                          |              |                      |
|              |                          |              | 1 : Yes              |
+--------------+--------------------------+--------------+----------------------+
| 10           | Alcohol intake           | alco         | Categorical - Binary |
|              |                          |              |                      |
|              |                          |              | 0 : No               |
|              |                          |              |                      |
|              |                          |              | 1 : Yes              |
+--------------+--------------------------+--------------+----------------------+
| 11           | Physical Activity        | active       | Categorical - Binary |
|              |                          |              |                      |
|              |                          |              | 0 : No               |
|              |                          |              |                      |
|              |                          |              | 1 : Yes              |
+--------------+--------------------------+--------------+----------------------+
| 12           | Cardio Disease           | cardio       | Categorical - Binary |
|              |                          |              |                      |
|              |                          |              | 0 : absent           |
|              |                          |              |                      |
|              |                          |              | 1 : present          |
+--------------+--------------------------+--------------+----------------------+

{{< pagebreak >}}

```{r setup, include=FALSE}
#| echo: false
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-library, include=FALSE}
#| echo: false
library(tidyverse)
library(patchwork)
library(kableExtra)
```

# Data Pre-processing

1.  Read data from .csv file.

```{r readData}
#| include: true
# Read CSV file.
rawCardioData = read_delim("data/cardio_train.csv", delim=";", 
                           show_col_types = FALSE) |>
                as_tibble()
```

```{r displayRawData}
#| echo: false
# Print Raw data
kbl(head(rawCardioData, 4), booktabs = T, align='c',
    caption="Raw Cardio Data") |>
  kable_styling(latex_options = c("striped","scale_down","HOLD_position"))
```

2.  Check missing values: There are no missing values in data.

```{r missingValues}
#| include: true
# Count total missing values in each column of data frame.
missing.values = rawCardioData |>
      pivot_longer(cols=1:ncol(rawCardioData), 
                   names_to = "Variables", 
                   values_to = "Value") |>
      group_by(Variables) |>
      summarise(missing.val = sum(is.na(Value)))|>
      pivot_wider(names_from=Variables, values_from=missing.val)
```

```{r print_NATable}
#| echo: false
# Print table Missing values per column
kbl(missing.values,booktabs = T, align='c',
    caption="Missing values per column") |>
  kable_styling(latex_options = c("striped","scale_down","HOLD_position"))
```

3.  Remove Duplicate rows from data. Total 24 duplicate rows are removed.

```{r duplicateRow}
#| include: true

# Remove duplicate rows using all variables except 'id'.
noUniqueData = rawCardioData |> 
              distinct(age,gender,height,weight,ap_hi,ap_lo,
                       cholesterol,gluc, smoke, alco, active, cardio,
                       .keep_all=TRUE)
```

4.  Convert category variables "*gender*", "*smoke*", "*cholesterol*", "*gluc*", "*alco*", "*active*", "*cardio*" to factor.

```{r factors}
#| echo: false

# Convert category variable to factor
tidyCardioData = noUniqueData |>
            mutate(
              gender = factor(gender, 
                             levels = c(1,2),
                             labels = c("Female", "Male")),
              cholesterol = factor(cholesterol,
                                  levels = c(1,2,3),
                                  labels = c("Normal",
                                             "Above Normal",
                                             "Well Above Normal")),
              gluc = factor(gluc,
                           levels = c(1,2,3),
                           labels = c("Normal",
                                      "Above Normal",
                                      "Well Above Normal")),
              smoke = factor(smoke, 
                            levels = c(0,1),
                            labels = c("No","Yes")),
              alco = factor(alco,
                           levels = c(0,1),
                           labels = c("No", "Yes")),
              active = factor(active,
                             levels = c(0,1),
                             labels = c("No", "Yes")),
              cardio = factor(cardio,
                             levels = c(1,0),
                             labels = c("Present", "Absent")) 
            )
```

5.  Convert scale of Age from days to year. Round it to whole number.

```{r ageScale}

# Convert scale of Age from days to year. Round it to whole number.
tidyCardioData = tidyCardioData |>
            mutate(age = round(age / 365))
```

6.  Display Tidy Data.

```{r displayTidyData}
#| echo: false
# Print Tidy Data
kbl(head(tidyCardioData, 4), booktabs = T, align='c',
    caption="Tidy Cardio Data") |>
  kable_styling(latex_options = c("striped","scale_down","HOLD_position"))
```

{{< pagebreak >}}

# Task 01

#### 1. Statistic Summary Table of continuous variables

```{r Task1_summaryTable}
#| echo: false
summaryTable = tidyCardioData |>
               select(age, height, weight, ap_hi, ap_lo) |>
               pivot_longer(cols=1:5, 
                            names_to="Variables", 
                            values_to="Values") |>
              group_by(Variables) |>
              summarise(Min = min(Values), 
                        Q1st = quantile(Values, p=0.25),
                        Median = median(Values),
                        Mean = mean(Values),
                        Q3rd = quantile(Values, p=0.75),
                        Max=max(Values) )

knitr::kable(summaryTable,
             format= "pipe", align = 'c', digits=2,
             col.names=c("Variable", "Min","1st qu.","Median",
                         "Mean","3rd qu.", "Max"),
             caption="Statistic summary of continuous variables")
```

#### 2. Plausibility check

##### Age :

Boxplot display there is outliers at lower age. Outlier age is 30. All data is collected for adults. This reason will keep data with age 30.

##### Height :

There are subject with height under 100cm. And subject's age is above 30 in data. It is may be incorrect data. As plausibility check, subject below 100cm height are removed.

##### Weight :

There are subject with weight below 25kg. And subject's age is above 30 in data. It is may be incorrect data. As plausibility check, subject below 25kg weight are removed.

##### Systolic blood pressure (SBP): ap_hi

From Statistic summary of Systolic blood pressure, there are outliers in dataset. There are negative values and very high values. Range between 20 and 380 will be consider as acceptable value.

Subject which has below 20 and above 380 SBP value are removed.

##### Diastolic blood pressure (DBP): ap_lo

From Statistic summary of Diastolic blood pressure, there are outliers in dataset. There are negative values and very high values. Range between 20 and 500 will be consider as acceptable value.

Subject which has below 20 and above 500 DBP value are removed.

```{r Task1_implausible_values}
#| include: true

# Filter implausible values from  height, weight, ap_hi and ap_lo.
# Height implausible values : below 100cm
# Weight implausible values : below 25kg
# ap_hi implausible values : below 20 and above 380
# ap_lo implausible values : below 20 and above 500
 cardioDataTask1 = tidyCardioData |>
              filter(height >= 100,
                     weight >= 25,
                     ap_hi > 20 & ap_hi <= 380,
                     ap_lo > 20 & ap_lo <= 500)
```

#### 3. Histogram and BoxPlot before & after remove implausible values.

```{r Task1-Age_plot}
#| echo: false

# Variable Age Box Plot
age_boxplot = ggplot(tidyCardioData) +
  geom_boxplot(aes(age)) +
  labs(x="age(years)")
  
# Variable Age Histogram
age_hist = ggplot(tidyCardioData) +
  geom_histogram(aes(x=age), fill='skyblue', color="black",
                 binwidth = 4) +
  #xlim(25,70) +
  labs(x="age(years)") +
  theme_classic()

# Variable Age Histogram and BoxPlot.
 age_hist + age_boxplot
```

```{r Task1_Height_plot}
#| echo: false

# Variable Height Box Plot
height_boxplot_before = ggplot(tidyCardioData) +
  geom_boxplot(aes(height))+
  labs(x="height(cm)", title = "Height Boxplot Before")

# Variable Height Histogram
height_hist_before = ggplot(tidyCardioData) +
  geom_histogram(aes(x=height), fill='skyblue', binwidth = 5) +
  labs(x="height(cm)", title = "Height Histogram Before")

# Variable Height density 
height_dens_before = ggplot(tidyCardioData) +
  geom_density(aes(x=height), fill='skyblue') + 
  labs(x="height(cm)", title = "Height Density Before")

 # Variable Height Box Plot after filter implausible values.
height_boxplot_after = ggplot(cardioDataTask1) +
  geom_boxplot(aes(height))+
  labs(x="height(cm)", title = "Height Boxplot After")

# Variable Height Histogram after filter implausible values.
height_hist_after = ggplot(cardioDataTask1) +
  geom_histogram(aes(x=height), fill='skyblue', bins = 100) +
  labs(x="height(cm)", title = "Height Histogram After")

# Variable Height density after filter implausible values.
height_dens_after = ggplot(cardioDataTask1) +
  geom_density(aes(x=height), fill='skyblue') + 
  labs(x="height(cm)", title = "Height Density After")

# Height histogram and BoxPlot before and after filter implausible values.
(height_boxplot_before + height_hist_before + height_dens_before) / (height_boxplot_after + height_hist_after  + height_dens_after )
```

```{r Task1_weight_plot}
#| echo: false

# Variable Weight Box-Plot 
weight_boxplot_before = ggplot(tidyCardioData) +
  geom_boxplot(aes(weight))+
  labs(x="weight(kg)", title = "Weight Boxplot Before")

# Variable Weight Histogram
weight_hist_before = ggplot(tidyCardioData) +
  geom_histogram(aes(x=weight), fill='skyblue', binwidth = 10) +
  labs(x="weight(kg)", title = "Weight Histogram Before")

# Variable Weight Density
weight_dens_before = ggplot(tidyCardioData) +
  geom_density(aes(x=weight), fill='skyblue') + 
  labs(x="weight(kg)", title = "Weight Density Before")
 
 # Variable Weight Box-Plot after filter implausible values.
weight_boxplot_after = ggplot(cardioDataTask1) +
  geom_boxplot(aes(weight))+
  labs(x="weight(kg)", title = "Weight Boxplot After")

# Variable Weight Histogram after filter implausible values.
weight_hist_after = ggplot(cardioDataTask1) +
  geom_histogram(aes(x=weight), fill='skyblue', binwidth = 10) +
  labs(x="weight(kg)", title = "Weight Histogram After")

# Variable Weight density after filter implausible values.
weight_dens_after = ggplot(cardioDataTask1) +
  geom_density(aes(x=weight), fill='skyblue') + 
  labs(x="weight(kg)", title = "Weight Density After")

# Weight histogram and BoxPlot before and after filter implausible values.
(weight_boxplot_before + weight_hist_before + weight_dens_before) / (weight_boxplot_after + weight_hist_after  + weight_dens_after )
```

```{r Task1_ap_hi_plot}
#| echo: false

# Variable SBP Box-Plot
sbp_boxplot_before = ggplot(tidyCardioData) +
  geom_boxplot(aes(ap_hi))+
  labs(x="Systolic BP", title = "SBP Boxplot Before")

# Variable SBP Histogram
sbp_hist_before = ggplot(tidyCardioData) +
  geom_histogram(aes(x=ap_hi), fill='skyblue', binwidth = 5) +
  labs(x="Systolic BP", title = "SBP Histogram Before")

# Variable SBP Density
sbp_dens_before = ggplot(tidyCardioData) +
  geom_density(aes(x=ap_hi), fill='skyblue') + 
  labs(x="Systolic BP", title = "SBP Density Before")

# Variable SBP Box-Plot after filter implausible values.
sbp_boxplot_after = ggplot(cardioDataTask1) +
  geom_boxplot(aes(ap_hi))+
  labs(x="Systolic BP", title = "SBP Boxplot After")

# Variable SBP Histogram after filter implausible values.
sbp_hist_after = ggplot(cardioDataTask1) +
  geom_histogram(aes(x=ap_hi), fill='skyblue', binwidth = 5) +
  labs(x="Systolic BP", title = "SBP Histogram After")

# Variable SBP Density after filter implausible values.
sbp_dens_after = ggplot(cardioDataTask1) +
  geom_density(aes(x=ap_hi), fill='skyblue') + 
  labs(x="Systolic BP", title = "SBP Density After")

# SBP histogram and BoxPlot before and after filter implausible values.
(sbp_boxplot_before + sbp_hist_before + sbp_dens_after) / 
  (sbp_boxplot_after + sbp_hist_after + sbp_dens_after)

```

```{r Task1_ap_lo_plot}
#| echo: false

# Variable DBP Box-Plot
dbp_boxplot_before = ggplot(tidyCardioData) +
  geom_boxplot(aes(ap_lo))+
  labs(x="Diastolic BP", title = "DBP Boxplot Before")

# Variable DBP Histogram
dbp_hist_before = ggplot(tidyCardioData) +
  geom_histogram(aes(x=ap_lo), fill='skyblue', binwidth = 5) +
  labs(x="Diastolic BP", title = "DBP Histogram Before")

# Variable DBP Density.
dbp_dens_before = ggplot(tidyCardioData) +
  geom_density(aes(x=ap_lo), fill='skyblue') + 
  labs(x="Diastolic BP", title = "DBP Density Before")

# Variable DBP Box-Plot
dbp_boxplot_after = ggplot(cardioDataTask1) +
  geom_boxplot(aes(ap_lo))+
  labs(x="Diastolic BP", title = "DBP Boxplot After")

# Variable DBP Histogram
dbp_hist_after = ggplot(cardioDataTask1) +
  geom_histogram(aes(x=ap_lo), fill='skyblue', binwidth = 5) +
  labs(x="Diastolic BP", title = "DBP Histogram After")

# Variable DBP Density after filter implausible values.
dbp_dens_after = ggplot(cardioDataTask1) +
  geom_density(aes(x=ap_lo), fill='skyblue') + 
  labs(x="Diastolic BP", title = "DBP Density After")

# DBP histogram and BoxPlot before and after filter implausible values.
(dbp_boxplot_before + dbp_hist_before + dbp_dens_after) / 
  (dbp_boxplot_after + dbp_hist_after + dbp_dens_after)
```

{{< pagebreak >}}

# Task 02

#### 1. Create new variable BMI (Body Mass Index).

$$
BodyMassIndex (kg/m^2) = Weight(kg)  /  [Height(m)]^2 
$$

```{r task2-1}
#| catch: true

# Create new variable BMI and save data as cardioDataTask2 
cardioDataTask2 = cardioDataTask1 |>
          mutate(BMI = weight / ((height/100)^2)) |> #Convert Height in meter
          relocate(BMI, .after=weight) # Move BMI Column after column Weight

```

```{r task2-1_printTable}
#| echo: false
#| tab-cap: Cardio data with BMI 
knitr::kable(head(cardioDataTask2, 4), digit=3, booktabs = T ) |>
             kable_styling(latex_options = c("HOLD_position","striped",
                                             "scale_down"))
```

#### 2. Summary table for the variable BMI for both cardio groups.

```{r task2-2}
#| echo: false
#| tbl-cap: BMI Summary for Cardio Group

# Create BMI Summary Table
bmi_summary = cardioDataTask2 |>
   group_by(cardio) |>
   summarise(min.BMI = min(BMI), 
             max.BMI = max(BMI), 
             mean.BMI = mean(BMI), 
             sd.BMI = sd(BMI), 
             median.BMI = median(BMI),
             Qu25.BMI = quantile(BMI, p=0.25),
             Qu75.BMI = quantile(BMI, p=0.75)
             )

```

```{r task2-2_tabledisplay}
#| echo: false
#| tab-cap: BMI Summary
knitr::kable(head(bmi_summary), format= "pipe", digit=2, align = 'c',  
             col.names = gsub("[.]", " ", names(bmi_summary)) )
```

{{< pagebreak >}}

# Task 03

#### 1. Correlation between the systolic blood pressure(SBP) and the BMI:

Spearman's rank correlation is computed to assess the correlation between the SBP and the BMI. Spearman Correlation is used because data is ordinal and there are still outlines present in data after plausibility check (Task1).

The Spearman Correlation coefficient is 0.28. The value is near 0, which means that there is a very low positive correlation between the SBP and the BMI.

```{r task3_1}
#| echo: false

# Correlation factor between systolic blood pressure(SBP) and BMI
cor_ap_hi_BMI = cor(cardioDataTask2$ap_hi, cardioDataTask2$BMI, 
                    method = "spearman")

cat("Correlation Coefficient with Spearman Method : ", 
    cor_ap_hi_BMI)

#Scatter Plot between SBP and BMI
BMI_aphi = ggplot(cardioDataTask2, aes(x=BMI, y=ap_hi, color=cardio)) +
           geom_point() + 
           labs(title = "Scatter Plot SBP ~ BMI")

BMI_aphi
```

{{< pagebreak >}}

#### 2. The SBP and the BMI correlation between the two classes of cardiovascular disease.

As per boxplot, subject with higher the SBP likely has cardiovascular disease. Also subject with higher BMI likely has cardiovascular disease.

```{r task3_2}
#| echo: false

# Filter data base on cardio 2 class i.e. Absent and Present.
cardioAbsent = cardioDataTask2 |>
               filter(cardio == "Absent") 

cardioPresent = cardioDataTask2 |>
               filter(cardio == "Present")

# Compare Scatter plot between SBP and BMI per cardio class.
sbp_BMI_cardioPresent = ggplot(cardioPresent,aes(x=BMI,y=ap_hi,color=cardio)) +
                           geom_point()  
                         
sbp_BMI_cardioAbsent = ggplot(cardioAbsent, 
                              aes(x=BMI, y=ap_hi,color=cardio)) +
                      geom_point() +
                      scale_color_manual(values=c("blue"))

sbp_BMI_cardioAbsent / sbp_BMI_cardioPresent +  
  plot_annotation(title = "Scatter plot : SBP ~ BMI per cardio class")

# Box Plot and density: cardio ~ SBP and cardio ~ BMI
task3_sbp_boxplot = ggplot(cardioDataTask2) +
                       geom_boxplot(aes(y=cardio, x=ap_hi, color=cardio)) +
                       theme_bw() 
task3_BMI_boxplot = ggplot(cardioDataTask2) +
                       geom_boxplot(aes(y=cardio, x=BMI, color=cardio)) +
                       theme_bw()
task3_sbp_density = ggplot(cardioDataTask2, 
                         aes(ap_hi, fill=cardio, color=cardio)) +
                  geom_density(alpha = 0.2)

task3_BMI_density = ggplot(cardioDataTask2, 
                         aes(BMI, fill=cardio, color=cardio)) +
                  geom_density(alpha = 0.2)
                      
(task3_sbp_boxplot + task3_sbp_density / task3_BMI_boxplot + task3_BMI_density ) + 
  plot_annotation(title = "Box & Denstity: SBP & BMI distribution in class of cardio")
```

{{< pagebreak >}}

# Task 04

#### 1. Correlation between the diastolic blood pressure (DBP) and the BMI:

As Task 3, Spearman Correlation is used to to assess the correlation between the DBP and the BMI.

The Spearman Correlation coefficient is 0.2476. The value is near 0, which means that there is a weak positive correlation between the DBP and the BMI.

```{r task4_1}
#| echo: false

# Correlation factor between diastolic blood pressure (DBP) and BMI
cor_ap_lo_BMI = cor(cardioDataTask2$ap_lo, 
                    cardioDataTask2$BMI, 
                    method = "spearman")

cat("Correlation Factor with Spearman Method : ", 
    cor_ap_lo_BMI)

#Scatter Plot between BMI and DBP
BMI_aplo = ggplot(cardioDataTask2, aes(x=BMI, y=ap_lo, color=cardio)) +
  geom_point() + 
  labs(title = "Scatter Plot DBP ~ BMI")

BMI_aplo
```

{{< pagebreak >}}

#### 2. The DBP and the BMI correlation between the two classes of cardiovascular disease.

As per boxplot, subject with higher the DBP likely has cardiovascular disease. Also subject with higher BMI likely has cardiovascular disease.

```{r task4_2}
#| echo: false

# Filter data base cardio 2 class i.e. Absent and Present.
cardioAbsent = cardioDataTask2 |>
               filter(cardio == "Absent") 

cardioPresent = cardioDataTask2 |>
               filter(cardio == "Present")

# Compare Scatter plot between BMI and DBP per cardio class.
aplo_BMI_cardioPresent = ggplot(cardioPresent,aes(x=BMI,y=ap_lo,color=cardio)) +
                           geom_point()
                         
aplo_BMI_cardioAbsent = ggplot(cardioAbsent, aes(x=BMI, y=ap_lo, color=cardio)) +                           geom_point() +
                          scale_color_manual(values=c("blue"))
                          
aplo_BMI_cardioAbsent / aplo_BMI_cardioPresent + 
  plot_annotation(title = "Scatter plot : DBP ~ BMI per cardio class")

# Box Plot : DBP ~ cardio and BMI ~ cardio
task3_ap_lo_boxplot = ggplot(cardioDataTask2) +
                       geom_boxplot(aes(y=cardio, x=ap_lo, color=cardio)) +
                       theme_bw() 
task3_BMI_boxplot = ggplot(cardioDataTask2) +
                       geom_boxplot(aes(y=cardio, x=BMI, color=cardio)) +
                       theme_bw() 

task3_ap_lo_boxplot / task3_BMI_boxplot + 
  plot_annotation(title = "Box plot : DBP and BMI distribution in the different categories of cardio")
```

{{< pagebreak >}}

# Task 05

Filter BMI, SBP and DBP data below 95% quantile threshold

```{r task5_filter}
# Remove subjects whose SBP, DBP and BMI the below 95% quantile threshold.
SBP_DBP_BMI_Q95 = cardioDataTask2 |>
                  filter(ap_hi <= quantile(ap_hi, 0.95),
                         ap_lo <= quantile(ap_lo, 0.95),
                         BMI <= quantile(BMI, 0.95))
```

#### 1. Correlation between SBP and BMI

```{r task5_ap_hi_1}
#| echo: false

# Correlation factor between SBP and BMI
cor_ap_hi_BMI1 = cor(SBP_DBP_BMI_Q95$ap_hi,
                     SBP_DBP_BMI_Q95$BMI, 
                     method = "spearman")

cat("Correlation Coefficient with Spearman Method : ", cor_ap_hi_BMI1)

#Scatter Plot between SBP and BMI
BMI_aphi_task5 = ggplot(SBP_DBP_BMI_Q95, aes(x=BMI, y=ap_hi, color=cardio)) +
                 geom_point() + 
                 labs(title = "Scatter Plot SBP ~ BMI")
BMI_aphi_task5
```

{{< pagebreak >}}

#### 2. Correlation between DBP and BMI

```{r task5_ap_lo_1}
#| echo: false

# Correlation factor between DBP and BMI
cor_ap_lo_BMI1 = cor(SBP_DBP_BMI_Q95$ap_lo,
                              SBP_DBP_BMI_Q95$BMI, 
                              method = "spearman")

cat("Correlation Coefficient with Spearman Method : ", cor_ap_lo_BMI1)

#Scatter Plot between SBP and BMI
BMI_aplo_task5 = ggplot(SBP_DBP_BMI_Q95, aes(x=BMI, y=ap_lo, color=cardio)) +
                 geom_point() + 
                 labs(title = "Scatter Plot SBP ~ BMI")
BMI_aplo_task5
```

{{< pagebreak >}}

#### 3. SBP, DBP and BMI correlation between cardio class

```{r task5_2}
#| echo: false

# Filter data base on cardio 2 class i.e. Absent and Present.
cardioAbsent1 = SBP_DBP_BMI_Q95 |>
               filter(cardio == "Absent") 

cardioPresent1 = SBP_DBP_BMI_Q95 |>
               filter(cardio == "Present")

# Compare Scatter plot between SBP and BMI per cardio class.
aphi_BMI_cardioPresent1 = ggplot(cardioPresent1,aes(x=BMI,y=ap_hi,color=cardio)) +
                           geom_point()  
                         
aphi_BMI_cardioAbsent1 = ggplot(cardioAbsent1, aes(x=BMI, y=ap_hi, color=cardio)) +                           geom_point() +
                          scale_color_manual(values=c("blue"))

aphi_BMI_cardioAbsent1 / aphi_BMI_cardioPresent1 +  
  plot_annotation(title = "Scatter plot : SBP ~ BMI per cardio class")

# Compare Scatter plot between DBP and BMI per cardio class.
aplo_BMI_cardioPresent1 = ggplot(cardioPresent1,aes(x=BMI,y=ap_lo,color=cardio)) +
                           geom_point()  
                         
aplo_BMI_cardioAbsent1 = ggplot(cardioAbsent1, aes(x=BMI, y=ap_lo, color=cardio)) +                           geom_point() +
                          scale_color_manual(values=c("blue"))

aplo_BMI_cardioAbsent1 / aplo_BMI_cardioPresent1 +  
  plot_annotation(title = "Scatter plot : DBP ~ BMI per cardio class")

# Box Plot : cardio ~ SBP, cardio ~ DBP and cardio ~ BMI
task5_aphi_boxplot = ggplot(SBP_DBP_BMI_Q95) +
                       geom_boxplot(aes(y=cardio, x=ap_hi, color=cardio)) +
                       theme_bw() 
task5_aplo_boxplot = ggplot(SBP_DBP_BMI_Q95) +
                       geom_boxplot(aes(y=cardio, x=ap_lo, color=cardio)) +
                       theme_bw() 
task5_BMI_boxplot = ggplot(SBP_DBP_BMI_Q95) +
                       geom_boxplot(aes(y=cardio, x=BMI, color=cardio)) +
                       theme_bw()
                      
task5_aphi_boxplot / task5_aplo_boxplot / task5_BMI_boxplot + 
  plot_annotation(title = "Box plot : SBP, DBP and BMI distribution in the different categories of cardio")
```

{{< pagebreak >}}

# Task 06

##### Analysis of Age distribution in the different categories of cardio**:**

```{r task6}
#| echo: false

# Age statistic summary Table group by cardio categories. 
age_cardio_Stat = tidyCardioData |>
               group_by(cardio) |>
               summarise(Min = min(age), 
                        Q1st = quantile(age, p=0.25),
                        Median = median(age),
                        Mean = mean(age),
                        Q3rd = quantile(age, p=0.75),
                        Max = max(age)
                        )

# Print Age summary Table
kbl(age_cardio_Stat, booktabs = T, digits=2, align='c',
       caption="Age Statistics Summary") |>
  kable_styling(latex_options = c("striped","HOLD_position"))

# Histogram stack bar with categories of cardio
task6_hist_fill = ggplot(cardioDataTask2) +
  geom_histogram(aes(age, fill=cardio),
                 color="black", binwidth=4, position="fill") +
  labs(x="age (years)")

# Boxplot of age
task6_boxplot = ggplot(cardioDataTask2) +
  geom_boxplot(aes(y=cardio, x=age, color=cardio)) +
  theme_bw() + 
  labs(x="age (years)")

# Density plot of age
task6_density = ggplot(cardioDataTask2, 
                         aes(age, fill=cardio, color=cardio)) +
                    geom_density(alpha = 0.3) +
                    labs(x="age (years)")

task6_hist_fill / task6_boxplot / task6_density +
  plot_layout(guides = 'collect') +
  plot_annotation(title = "Age Distribution in outcome variable cardio")

```

##### Conclusion:

Median age for cardiac disease present is higher than for cardiac disease absent by 4 years.

50% of subjects with cardiac disease present are in age between 50 and 60. 50% of subjects with cardiac disease absent are in age between 46 and 56. Age distribution for cardiac disease present is left-skewed. Age is spread in same range for cardiac disease absent and present.

Histogram shows, present of cardiac disease is increase with age from 25% to 75%.

{{< pagebreak >}}

# Task 07

#### **Task Description:**

Create a plot that show the distribution of age for both types of gender and both types of cardio.

```{r task7-plot}
# Boxplot 
ggplot(cardioDataTask2) +
  geom_boxplot(aes(x=cardio, y=age, color=cardio)) +
  facet_grid( ~ gender) +
  ylab("age (year)") +
  theme_bw()
```

{{< pagebreak >}}

# Task 08

#### **Task Description:**

Extend this plot by taking the different types of glucose into account.

```{r task8}

ggplot(cardioDataTask2) +
  geom_boxplot(aes(x=cardio, y=age, color=cardio)) +
  facet_grid(gender ~ gluc) +
  theme_bw()
```

{{< pagebreak >}}

# Task 09

#### 1. Contingency table :

Risk parameters i.e. smoking, alcohol, and insufficient physical activity distribution between two types of cardio.

```{r task9-table}
#| echo: false

# Freq. table using ftable()
# Select only Catagory variables 3 risk factors and cardio
task9Data = cardioDataTask2 |>
            select(active, smoke, alco, cardio)

# Covert ftable to dataframe to print with kabelExtra.
overviewTableF = as.data.frame(ftable(task9Data)) |>
                 pivot_wider(names_from = cardio, values_from = Freq) |>
                 rename(Active=active, Smoke=smoke, Alcohol=alco)

# Print Contingency Table                
kbl(overviewTableF,  booktabs = T,
    caption="Contingency Table: Distributions of three risk factors") |>
    kable_styling(latex_options = c("striped", "HOLD_position")) |>
    add_header_above(c(" "," "," ","Cardio" = 2))
```

#### 2. Chi-squared test:

Statistical test to test if the relationship of two variables is independent or not.

-   Chi-squared test : Cardio \~ Alcohol

```{r task9-x2-alco}
#| echo: false

# chi-test bwtween catagory variables cardio and alcohol
chiTest_cardio_alco = chisq.test(cardioDataTask2$cardio,
                                   cardioDataTask2$alco)
chiTest_cardio_alco
```

-   Chi-squared test : Cardio \~ Smoking

```{r task9-x2-smoke}
#| echo: false

# chi-test bwtween catagory variables cardio and smoking
chiTest_cardio_smoke = chisq.test(cardioDataTask2$cardio,
                                   cardioDataTask2$smoke)
chiTest_cardio_smoke
```

-   Chi-squared test : Cardio \~ Activity

```{r task9-x2-active}
#| echo: false

# chi-test bwtween catagory variables cardio and physical activity
chiTest_cardio_active = chisq.test(cardioDataTask2$cardio,
                                   cardioDataTask2$active)
chiTest_cardio_active
# summary(table(cardioDataTask2$cardio,cardioDataTask2$active))
```

#### 3. Conclusion:

-   Assumption that independent between variables 'cardio' and 'Alcohol' is rejected, as statistical test p-value is less than 0.05%

-   Assumption that independent between variables 'cardio' and 'Smoking' is rejected, as statistical test p-value is less than 0.05%

-   Assumption that independent between variables 'cardio' and 'Physical Activity' is rejected, as statistical test p-value is less than 0.05%
